---
title: "Task 2: Mini exploratory report on juvenile showshoe hares in Bonanza Creek Experimental Forest"
author: "By: Claire Gonzales"
date: "November 10, 2021"
output: 
  html_document: 
    theme: lumen
---
## 1. Project Overview

Bonanza Creek LTER is part of the US Long Term Ecological Research Network. See more information about the site here. In this “mini exploratory report,” you will explore the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012. 



## 2. Data and Analysis

Full metadata can be found at the following URL: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-bnz.55.22

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# install packages

library(tidyverse)
library(janitor)
library(here)
library(lubridate) # for `ymd()` function
library(broom)
library(ggbeeswarm)
library(kableExtra)
library(effsize)

```


```{r}
# Read in raw csv file
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
```

## 3. Initial Juvenile Hare Investigation

### 3a. Annual juvenile hare trap counts
  
```{r}
# Count the total number of juvenile hare trappings during each year of the study
hare_juv_year <- bonanza_hares %>% 
  filter(age == "j") %>% # only include observations of juvenile hares
  mutate(date = mdy(date)) %>% # lubridate() to transform date from myd forat to YYYY-MM-DD format
  mutate(year = year(date)) %>% # add a column with just the year
  group_by(year) %>% # group by year
  summarize(count = n()) # count of observations per year

```

```{r}
# Create a finalized data visualization of the counts by year.

ggplot(data = hare_juv_year, aes(x = year, y = count)) + # setting up ggplot
  geom_bar(stat = "identity", fill = "steelblue3") + # color coding for year
  labs(x = "Hare count", # axis and main labels
       y = "Year",
       fill = none) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) # centering title

```

**Figure 1.** Number of juvenile snowshoe hares that were captured within the Bonanza Creek Experimental Forest between the years 1999 and 2012. (Data: Kielland et. al., 2017)

```{r, include = FALSE}
# Summary statistics for in-line description
hare_juv_summary <- bonanza_hares %>% # running a summary statistics from original data
  filter(age == "j") %>%
  mutate(date = mdy(date)) %>% # updating date format
  mutate(year = year(date)) %>% # include a column with year only
  group_by(year) %>% 
  summarize(count = n()) %>% # total observations by year
  summarize(mean_count = round(mean(count, na.rm = TRUE), 1), # create a table based on those counts of observations per year
            median_count = median(count, na.rm = TRUE),
            max_count = max(count, na.rm = TRUE),
            min_count = min(count, na.rm = TRUE),
            sd_count = round(sd(count, na.rm = TRUE), 1),
            total = sum(count))

```


The capture data above (**Figure 1**) reflect that the amount of juvenile snowshoe hares captured was highest in the year 1999 (n = `r hare_juv_summary$max_count`) and steadily decreased over time. In total, `r hare_juv_summary$total` juvenile snowshoe hares were captured throughout the project duration (1999 - 2012) and the mean yearly capture was `r hare_juv_summary$max_count` $\pm$ `r hare_juv_summary$sd_count` (mean $\pm$ sd). However, it is important to note that this capture data is dependent on how many traps are set out. To account for this, future projects should set out the same amount of traps each year or list results as a percentage of number of hares captured per number of traps deployed annually. 

### 3b. Juvenile hare weight visualization


```{r}
hare_juv_weight <- bonanza_hares %>% 
  filter(age == "j") %>%
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(grid = case_when( # updating site names (column grid) to have capitalization
    grid == "bonbs" ~ "Bonbs",
    grid == "bonmat" ~ "Bonmat",
    grid == "bonrip" ~ "Bonrip")) %>% 
  mutate(sex = case_when(
    sex == "f" ~ "Female",
    sex == "m" ~ "Male",
    sex == "NA" ~ "N/A"))


ggplot(data = hare_juv_weight, aes(x = grid, y = weight)) + # beeswarm plot with box plot , facet wrapped by sex
  geom_beeswarm(aes(color = grid)) +
  scale_color_manual(values = c("goldenrod2","cyan4","coral")) +
  geom_boxplot(fill = NA, width = 0.3, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=3, 
               color="black", 
               fill="black") +
  facet_grid(~sex) +
  labs(x = "Site",
       y = "Body mass (g)",
       color = "Site") +
  theme_minimal() +
  theme(panel.spacing = unit(2, "lines"))

```

**Figure 2.** Comparisons of body mass values (grams) juvenile snowshoe hares captured at Bonanza Creek Experimental Forest between the years 1999 and 2012. Female snowshoe hares are in left pane, followed by male snowshoe hares (middle pane) and hares for which sex was not collected (right pane). Each pane contains body mass measurements for juvenile snowshoe hares captures at each site (Bonbs, Bonmat, and Bonrip). Capture observations from each site are further described with a box plot. The box's middle black line represents the median values, while the outsides of the box represent the 25th and 75th percentile values. The black dot represents the mean value of observations. (Data: Kielland et. al., 2017)

  
The data above (**Figure 2**) reflects that the body mass (g) measurements of male juvenile snowshoe hares had slightly larger mean and median values compared to those measurements made for female juvenile snowshoe hares. In general, these data contain very few outliers, with only one positive outlier for female observations at the Bonrip site and one negative outlier for male observations at the Bonbs site. Notably, the heaviest snowshoe hares for both male and female measurements were captured at the Bonrip site. Conversely, the highest mean and median values were in the Bonbs site across sexes. Few hares could not be sexed and all of those observations (NA; right pane) were constrained to Bonmat and Bonrip sites.

### 3c. Juvenile weight comparison (male & female snowshoe hares)


```{r, include=FALSE}
# statistics for in-line description

f_hare_juve_weight <- hare_juv_weight %>% # creating a vecctor for female body mass values
  filter(sex == "Female") %>% 
  pull(weight) # the pull() part makes a vector of the values

m_hare_juve_weight <- hare_juv_weight %>% # creating a vector for male body mass values
  filter(sex == "Male") %>% 
  pull(weight)


# Exploratory histograms and QQ plots to check assumptions
hist(f_hare_juve_weight) 
# Looks normal

qqnorm(f_hare_juve_weight) 
# Looks linear

hist(m_hare_juve_weight) 
# A little skewed negatively

qqnorm(m_hare_juve_weight)
# Looks fairly linear

# Based on these exploratory tests, we can assume normality of the data and conduct a t test

hare_juv_weigth_ttest <- t.test(f_hare_juve_weight, m_hare_juve_weight)

hare_juv_weigth_ttest_tidy <- tidy(hare_juv_weigth_ttest) #make it tidy
hare_juv_weigth_ttest_tidy # call up output


# other summary stats needed

mean_m <- round(mean(m_hare_juve_weight),2)
mean_f <- round(mean(f_hare_juve_weight, na.rm = TRUE),2)
median_m <- round(median(m_hare_juve_weight),2)
median_f <- round(median(f_hare_juve_weight, na.rm = TRUE),2)
sd_m <- round(sd(m_hare_juve_weight),2)
sd_f <- round(sd(f_hare_juve_weight, na.rm = TRUE),2)

mean_diff <- mean_m - mean_f

# calculate effect size
juv_hare_effsize <- cohen.d(f_hare_juve_weight, m_hare_juve_weight, na.rm = TRUE)

juv_hare_effsize # call up output
```

```{r}
# create table of important values

hare_juv_compare <- hare_juv_weight %>% 
  group_by(sex) %>% 
  summarize(
      Mean = round(mean(weight, na.rm = TRUE),1),
      SD = round(sd(weight, na.rm = TRUE),1),
      Count = n())

kable(hare_juv_compare,
      align = "c",
      booktabs = TRUE,
      col.names = c("Sex", # labeling column names with unit
                    "Mean (g)",
                    "Standerd Deviation (g)",
                    "Count"),
      caption = "**Table 1** Shows a summary statistics table for body weight distribution with respect to sex of juvenile snowshoe hares captured in Bonanza Creek Experimental Forest between the years 1999 and 2012. (Data: Kielland et. al., 2017)") %>% 
   kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

The body weight (g) distribution of female (n = `r hare_juv_compare$Count[1]`) and male (n = `r hare_juv_compare$Count[2]`) juvenile snowshoe hares depicted in **Table 1**. The difference in means between weights of male and female snowshoe hares is `r mean_diff` (g), with a small effect size (Cohen's *d* = `r juv_hare_effsize[3]`). Despite this, the difference in means is significant (Welch's two-sample t-test: t(`r round(hare_juv_weigth_ttest_tidy$parameter,2)` = `r round(hare_juv_weigth_ttest_tidy$estimate,2)`, p = `r round(hare_juv_weigth_ttest_tidy$p.value,3)`).

### 3d. Relationship between juvenile weight & hind foot length

```{r}
ggplot(data = hare_juv_weight, aes(x = weight, y = hindft)) + #Initial visualization of hind foot length and weight of hare
  geom_point() +
  labs( x = "Body mass (g)" ,
        y = "Hind foot length (mm)") +
  theme_minimal() # figure 3
```

**Figure 3** Visualization of initial relationship between body mass (g) with hind leg length (mm) of juvenile snowshow hares captured in Bonanza Creek Experimental Forest between the years 1999 and 2012. (Data: Kielland et. al., 2017)

```{r, include = FALSE}
# Assessing liner regression to describe how juvenile snowshoe hare weight changes with respect to hind foot length.
#First, explore assumptions

# Creating vectors
juv_weight <- hare_juv_weight %>% 
  pull(weight)
juv_hindft <-hare_juv_weight %>% 
  pull(hindft)

#test normality
hist(juv_weight) #pretty normal
qqnorm(juv_weight) #pretty straight
hist(juv_hindft) #less normal
qqnorm(juv_hindft) #somewhat straight

# running linear model
juvenile_lm <- lm(hindft ~ weight, data = hare_juv_weight)
juvenile_lm # call up value

#tidy and glance versions of linear regression
juvenile_lm_tidy <- tidy(juvenile_lm)
juvenile_lm_tidy # call up value

juvenile_lm_glance <- glance(juvenile_lm)
juvenile_lm_glance # call up value

# plot(juvenile_lm)
# (formatting so this doesnt show up in knitted doc)

# pearson's r
juvenile_cor <- cor.test(hare_juv_weight$weight, hare_juv_weight$hindft)
juvenile_cor # call up value

#tidy 
juvenile_cor_tidy <- tidy(juvenile_cor)
juvenile_cor_tidy
```

The relationship between snowshoe hare body mass (g) and hind foot length (mm) is fairly linear across sex and site (**Figure 3.**). In exploring this more, the simple linear regression reflected that hare body mass loosely predicts length of hind foot (p < 0.001, R^2^ = `r round(juvenile_lm_glance$r.squared,3)`). If all other factors remain constant, on average we would expect hind foot length to increase by `r round(juvenile_lm_tidy$estimate[2],3)`mm per 1g increase in body mass (slope = `r round(juvenile_lm_tidy$estimate[2],3)`mm/g). Body mass (g) and hind foot length (mm) of juvenile snowshoe hares have medium positive and significant coorelation (Pearson's *r* = `r round(juvenile_cor_tidy$estimate,2)`, p < 0.001). Lastly, diagnostic plots (not included) revealed normally distributed residuals, but residuals were fairly heteroscedastic. This is a concern in using this linear regression to draw conclusions about the population.

## 4. Summary of Results

This initial exploration of the data reflected the following results:


## 5. Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14
