---
title: "Task 2: Mini 'exploratory report' on juvenile snowshoe hares in Bonanza Creek Experimental Forest"
author: "By: Claire Gonzales"
date: "11/6/2021"
output: 
  html_document: 
    theme: simplex
---

## 1. Project Overview

Bonanza Creek LTER is part of the US Long Term Ecological Research Network. See more information about the site here. In this “mini exploratory report,” you will explore the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012. The data citation and link to metadata (including variable definitions) are below.


Data citation: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

Link to metadata: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-bnz.55.22


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install packages

library(tidyverse)
library(janitor)
library(here)

```

# Read in Data

```{r}
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
```

